Crystal as a thermodynamic computer

### 2024-09-05 paper/main.tex
update
update depth label

With the proposed Ising model, we are interested in how efficiently we can drive the system to the ground state, which encodes the output of the CA.
Since the natural cooling process is highly nontrivial to model, we consider a theoretical model for cooling, which is simulated annealing (SA)~\cite{Kirkpatrick1983}.
\jinguo{random fluctuations may be linear in time considering the parallelism of the system.}
It is known that when the system and heat bath are at thermal equilibrium, 
the probability of a spin configuration $|\mathbf{s}\rangle$ is given by the Boltzmann distribution:

\jinguo{please do not use the ket bra notation}
\begin{equation}
    |\mathbf{s}\rangle = \frac{1}{Z}\exp\left(- \frac{E_{|\mathbf{s}\rangle}}{k_bT}\right)
\end{equation}

where $Z = \sum_{|\mathbf{s}\rangle} \exp\left(-\frac{E_{|\mathbf{s}\rangle}}{k_bT}\right)$ is the partition function, 
$E_{|\mathbf{s}\rangle}$ is the energy of the each of $2^n$ spin configuration, labeled by $|\mathbf{s}\rangle = |s_1, s_2, \cdots, s_n\rangle$, $k_b$ is the Boltzmann constant, and $T$ is the temperature of the heat bath. 
In the following, we set $k_b = 1$.

We employ the Metropolis-Hastings algorithm~\cite{metropolis1953equation} to characterize the thermodynamic process of the TC system as modulated by the heat bath.
This algorithm perform a random walk in the state space of the TC, each step we try to flip a spin,
and the probability of accepting a move from state $|\mathbf{s}\rangle$ to $|\mathbf{s}^\prime\rangle$ is given by:

\begin{equation}
    P_{\text{accept}} = \frac{1}{1+\exp\left(\frac{E_{|\mathbf{s}^\prime\rangle} - E_{|\mathbf{s}\rangle}}{T}\right)}
\end{equation}

@@

More specifically, the computation, progressing from the input layer to the output layer, involves the following steps: (1) Initialize the input layer configuration by adjusting the onsite energy for each spin. (2) Determine the number of layers $W$ that will interact with the external heat bath in each heating-annealing process. Group the layers in sets of $W$ from input to output, labeling them as groups $1, 2, \cdots, \frac{T}{W}$. The number of heating-annealing repetitions will then be $\frac{T}{W}$, where T is the total number of layers. (3) Perform the heating-annealing process on each group sequentially.

@@

Assume there is an annealing procedure that may prevent the layers in the Heat Zone from stabilizing in the local ground state, or equivalently, that an error occurs in the computation process with a probability $\epsilon$. Repeating this procedure $\frac{T}{W}$ times would result in a probability of $(1-\epsilon)^{\frac{T}{W}}$ to successfully find the proper global ground state. To maintain a constant success probability as $T \rightarrow +\infty$, the error probability $\epsilon$ should scale as $\frac{W}{T}$.

@@

\begin{equation}
    \left( \frac{\lambda_2}{\lambda_1} \right) ^ {t_{th}} = \epsilon \rightarrow t_{th} \sim \ln(\frac{T}{W})e^{W}
\end{equation}

And the total annealing time scale as $\frac{T}{W}\ln(\frac{T}{W})e^W$.

@@

The block annealing process then can be reconstruct as slowly moving
the middle of the heat bath from one end to the other. Defining $\Delta E_{max}$ to be the
largest possible energy change when a spin is fliped.
The effective heat zone should then be $[y-W, y]$, where position $y$ satisfying 
$\exp(\frac{\Delta E_{max}}{T(y)}) = 0.1$ and $y<middle(t)$. Since the layers
in fronts of the effective heat zone can only be fliped with a $\epsilon$ probability if the local energy would increase. 
The effective width $W$ then should scale as $\exp(-\Delta E_{max} \Lambda^{cW}) \sim \epsilon \sim \frac{W}{T}$, where $c$ is the empirical constant for effective width and can be determined during the experiment.

@@

\begin{equation}
    \label{eq:total-annealing-time}
    t_{total} = \frac{T}{W}(\ln(\frac{T}{W}))^{1+ \frac{1}{c\ln(\Lambda)}}
\end{equation}

@@

We fixed the number of layers and denote $n$ as the width of each layer in this material.

However, we fail to derive an explicitly expression for the time complexity in this direction. 
Numerical result shows that computing complexity is linear to $n$. Thus a total time
for the computation should scale as $nm\log^{1 + \frac{1}{c\ln(\Lambda)}}(m)$

@@

Denote $S$ to be the length of the material along the ``Space'' axis, $T$
to be the length of the material along the ``Time'' axis. 
Let us define a simpler but equivalent energy model on a 2D lattice of size $S \times T$, whose hamiltonian
is given by
\begin{equation}\label{eq:toy-hamiltonian}
H = \sum_{t=1}^{T-1}\sum_{i=1}^{S} (n_{i+1, t} \land \neg(n_{i, t} \land \neg n_{i-1, t})) \veebar n_{i,t+1},
\end{equation}

@@

\jinguo{show a comparison from right to left}\jinguo{I have already used D as the dimension}}
\ym{Great, i will do this}
Here we conducted numerical experiments to examine both the time complexity along the ``Space'' direction and the ``Time'' direction.
We use binary search to determine the speed of the slowly moving heat bath at which
the simulated annealing procedure would produce 
a ground state configuration with $50\%$ probability. Denote this speed 
as $v$, the simulated annealing procedure will then take $\frac{T}{v}$
stages to complete the computation, each stage we try to flip a spin with 
probability define in (\ref{eq:non-equilibrium-SA}). We name $\frac{T}{v}$ 
as Sweep Time.

@@

[Fig. \ref{fig:main-numerical-result}(a)] shows the relationship between
$\ln(\ln(T))$ and $\ln(\frac{\text{Sweep Time}}{T})$ shoule be linear, which implies 
the time complexity along the ``Time'' direction should scale as $T\ln(T)^{f(\Lambda)}$. 
We then fit the slopes of each line and $\frac{1}{\ln(\Lambda)}$ and find the 
empirical constant $c$ in (\ref{eq:total-annealing-time}) to be $3$. [Fig. \ref{fig:main-numerical-result}(b)] shows 
the the relationship between Sweep Time and $S$ should be linear. Combining these two results, 
we can conclude that the time complexity should scales as $TS\log(T)^{1+\frac{1}{3\ln(\Lambda)}}$, 
which is consistent with our theoretical prediction in (\ref{eq:total-annealing-time}).
% \begin{figure}[h]
%     \centering
%     \begin{minipage}[b]{0.485\columnwidth}
%         \centering
%         \includegraphics[width=\columnwidth,keepaspectratio]{../notes/images/TimeComplexityalongTaxis.png}
%         \caption{Time v.s. Depth of the material}
%         \label{fig:image1}
%     \end{minipage}
%     \hspace{0.001\columnwidth} % 两张图片之间的水平间距
%     \begin{minipage}[b]{0.485\columnwidth}
%         \centering
%         \includegraphics[width=\columnwidth,keepaspectratio]{../notes/images/toymodel_time_vs_width.png}
%         \caption{Time v.s. width of the material}
%         \label{fig:image2}
%     \end{minipage}
% \end{figure}

>>>>

With the proposed Ising model, we are interested in how efficiently we can drive the system to the ground state, which encodes the output of the CA.
Since the natural cooling process is highly nontrivial to model, we consider a theoretical model for cooling, which is simulated annealing (SA)~\cite{Kirkpatrick1983}.
\jinguo{random fluctuations may be linear in time considering the parallelism of the system.}
It is known that when the system and heat bath are at thermal equilibrium, 
the probability of a spin configuration $\mathbf{s}$ is given by the Boltzmann distribution:

\jinguo{please do not use the ket bra notation}
\begin{equation}
    \mathbf{s} = \frac{1}{Z}\exp\left(- \frac{E_{\mathbf{s}}}{k_bT}\right)
\end{equation}

where $Z = \sum_{\mathbf{s}} \exp\left(-\frac{E_{\mathbf{s}}}{k_bT}\right)$ is the partition function, 
$E_{\mathbf{s}}$ is the energy of the each of $2^n$ spin configuration, labeled by $\mathbf{s} = (s_1, s_2, \cdots, s_n)$, $k_b$ is the Boltzmann constant, and $T$ is the temperature of the heat bath. 
In the following, we set $k_b = 1$.

We employ the Metropolis-Hastings algorithm~\cite{metropolis1953equation} to characterize the thermodynamic process of the TC system as modulated by the heat bath.
This algorithm perform a random walk in the state space of the TC, each step we try to flip a spin,
and the probability of accepting a move from state $\mathbf{s}$ to $\mathbf{s}^\prime$ is given by:

\begin{equation}
    P_{\text{accept}} = \frac{1}{1+\exp\left(\frac{E_{\mathbf{s}^\prime} - E_{\mathbf{s}}}{T}\right)}
\end{equation}

@@

More specifically, the computation, progressing from the input layer to the output layer, involves the following steps: (1) Initialize the input layer configuration by adjusting the onsite energy for each spin. (2) Determine the number of layers $W$ that will interact with the external heat bath in each heating-annealing process. Group the layers in sets of $W$ from input to output, labeling them as groups $1, 2, \cdots, \frac{D}{W}$. The number of heating-annealing repetitions will then be $\frac{D}{W}$, where D is the total number of layers. (3) Perform the heating-annealing process on each group sequentially.

@@

Assume there is an annealing procedure that may prevent the layers in the Heat Zone from stabilizing in the local ground state, or equivalently, that an error occurs in the computation process with a probability $\epsilon$. Repeating this procedure $\frac{D}{W}$ times would result in a probability of $(1-\epsilon)^{\frac{D}{W}}$ to successfully find the proper global ground state. To maintain a constant success probability as $D \rightarrow +\infty$, the error probability $\epsilon$ should scale as $\frac{W}{D}$.

@@

\begin{equation}
    \left( \frac{\lambda_2}{\lambda_1} \right) ^ {t_{th}} = \epsilon \rightarrow t_{th} \sim \ln(\frac{D}{W})e^{W}
\end{equation}

And the total annealing time scale as $\frac{D}{W}\ln(\frac{D}{W})e^W$.

@@

The block annealing process then can be reconstruct as slowly moving
the middle of the heat bath from one end to the other. Defining $\Delta E_{max}$ to be the
largest possible energy change when a spin is fliped.
The effective heat zone should then be $[y-W, y]$, where position $y$ satisfying 
$\exp(\frac{\Delta E_{max}}{T(y)}) = 0.1$ and $y<middle(t)$. Since the layers
in fronts of the effective heat zone can only be fliped with a $\epsilon$ probability if the local energy would increase. 
The effective width $W$ then should scale as $\exp(-\Delta E_{max} \Lambda^{cW}) \sim \epsilon \sim \frac{W}{D}$, where $c$ is the empirical constant for effective width and can be determined during the experiment.

@@

\begin{equation}
    \label{eq:total-annealing-time}
    t_{total} = \frac{D}{W}(\ln(\frac{D}{W}))^{1+ \frac{1}{c\ln(\Lambda)}}
\end{equation}

@@

We fixed the number of layers and denote $S$ as the width of each layer in this material.

However, we fail to derive an explicitly expression for the time complexity in this direction. 
Numerical result shows that computing complexity is linear to $S$. Thus a total time
for the computation should scale as $DS\log^{1 + \frac{1}{c\ln(\Lambda)}}(D)$

@@

Denote $S$ to be the length of the material along the ``Space'' axis, $D$
to be the length of the material along the ``Time'' axis. 
Let us define a simpler but equivalent energy model on a 2D lattice of size $S \times D$, whose hamiltonian
is given by
\begin{equation}\label{eq:toy-hamiltonian}
H = \sum_{t=1}^{D-1}\sum_{i=1}^{S} (n_{i+1, t} \land \neg(n_{i, t} \land \neg n_{i-1, t})) \veebar n_{i,t+1},
\end{equation}

@@

\jinguo{show a comparison from right to left}\jinguo{I have already used D as the dimension}
\ym{Great, i will do this}
Here we conducted numerical experiments to examine both the time complexity along the ``Space'' direction and the ``Time'' direction.
We use binary search to determine the speed of the slowly moving heat bath at which
the simulated annealing procedure would produce 
a ground state configuration with $50\%$ probability. Denote this speed 
as $v$, the simulated annealing procedure will then take $\frac{D}{v}$
stages to complete the computation, each stage we try to flip a spin with 
probability define in (\ref{eq:non-equilibrium-SA}). We name $\frac{D}{v}$ 
as Sweep Time.

@@

[Fig. \ref{fig:main-numerical-result}(a)] shows the relationship between
$\ln(\ln(D))$ and $\ln(\frac{\text{Sweep Time}}{D})$ shoule be linear, which implies 
the time complexity along the ``Time'' direction should scale as $D\ln(D)^{f(\Lambda)}$. 
We then fit the slopes of each line and $\frac{1}{\ln(\Lambda)}$ and find the 
empirical constant $c$ in (\ref{eq:total-annealing-time}) to be $3$. [Fig. \ref{fig:main-numerical-result}(b)] shows 
the the relationship between Sweep Time and $S$ should be linear. Combining these two results, 
we can conclude that the time complexity should scales as $DS\log(D)^{1+\frac{1}{3\ln(\Lambda)}}$, 
which is consistent with our theoretical prediction in (\ref{eq:total-annealing-time}).
% \begin{figure}[h]
%     \centering
%     \begin{minipage}[b]{0.485\columnwidth}
%         \centering
%         \includegraphics[width=\columnwidth,keepaspectratio]{../notes/images/TimeComplexityalongTaxis.png}
%         \caption{Time v.s. Depth of the material}
%         \label{fig:image1}
%     \end{minipage}
%     \hspace{0.001\columnwidth} % 两张图片之间的水平间距
%     \begin{minipage}[b]{0.485\columnwidth}
%         \centering
%         \includegraphics[width=\columnwidth,keepaspectratio]{../notes/images/toymodel_time_vs_width.png}
%         \caption{Time v.s. width of the material}
%         \label{fig:image2}
%     \end{minipage}
% \end{figure}
